---
title: "Final project"
author: "Anisha Ganeshkumar"
date: "01/12/2019"
output: html_document
---

```{r}
library(dplyr)
library(ggplot2)
library(stringr)
```
```{r}
setwd("~/Datasets/")
airbnb <- read.csv('airbnb_listings.csv', sep = ';', comment.char = '#', stringsAsFactors = FALSE)

columns = c('ID', 'Host.Name','Host.Since','Host.Acceptance.Rate','Host.Listings.Count','City','State','Market','Country.Code','Country','Latitude','Longitude','Property.Type',
'Room.Type','Accommodates','Bathrooms','Bedrooms','Amenities','Price','Weekly.Price','Monthly.Price','Security.Deposit',
'Cleaning.Fee','Number.of.Reviews','Extra.People','Guests.Included','Host.Verifications','Property.Type','Cancellation.Policy','Review.Scores.Rating',
'Review.Scores.Accuracy','Review.Scores.Checkin','Review.Scores.Cleanliness','Review.Scores.Communication','Review.Scores.Location','Review.Scores.Value',
'Reviews.per.Month','Features')

airbnb <- airbnb %>% select(columns)

```
Data Cleaning

```{r}
#create a function to remove the special charachters from a string
remove_special_charachters <- function(x) {
  return(gsub("[^-0-9/.]+", "", x))
}
```

```{r}
# first we will get rid of records with false value (alphabetical value)
airbnb <- airbnb[!grepl('[[:alpha:] ]', airbnb$Host.Acceptance.Rate),, drop = FALSE]
# Then we remove the "%" from each record 
airbnb$Host.Acceptance.Rate <- sapply(airbnb$Host.Acceptance.Rate, remove_special_charachters)
# Convert the column to a numerical column
airbnb$Host.Acceptance.Rate<-as.numeric(airbnb$Host.Acceptance.Rate)
# We convert the column to numerical columns
airbnb$Host.Listings.Count<-as.numeric(airbnb$Host.Listings.Count)
#Convert column to numerical column
airbnb$Accommodates <-as.numeric(airbnb$Accommodates)
#Convert column to numeric column
airbnb$Bathrooms <- as.numeric(airbnb$Bathrooms)
#Convert column to numeric
airbnb$Bedrooms <- as.numeric(airbnb$Bedrooms)
```

```{r}
airbnb$Price <- as.numeric(airbnb$Price)
airbnb$Weekly.Price <- as.numeric(airbnb$Weekly.Price)
airbnb$Monthly.Price <- as.numeric(airbnb$Monthly.Price)
airbnb$Security.Deposit <- as.numeric(airbnb$Security.Deposit)
airbnb$Cleaning.Fee <- as.numeric(airbnb$Cleaning.Fee)
airbnb$Number.of.Reviews <- as.numeric(airbnb$Number.of.Reviews)
airbnb$Extra.People <- as.numeric(airbnb$Extra.People)
airbnb$Guests.Included <- as.numeric(airbnb$Guests.Included)
airbnb$Review.Scores.Rating <- as.numeric(airbnb$Review.Scores.Rating)
airbnb$Review.Scores.Accuracy <- as.numeric(airbnb$Review.Scores.Accuracy)
airbnb$Review.Scores.Checkin <- as.numeric(airbnb$Review.Scores.Checkin)
airbnb$Review.Scores.Cleanliness <- as.numeric(airbnb$Review.Scores.Cleanliness)
airbnb$Review.Scores.Communication <- as.numeric(airbnb$Review.Scores.Communication)
airbnb$Review.Scores.Location <- as.numeric(airbnb$Review.Scores.Location)
airbnb$Review.Scores.Value <- as.numeric(airbnb$Review.Scores.Value)
airbnb$Reviews.per.Month <- as.numeric(airbnb$Reviews.per.Month)
```
```{r}
str(airbnb)
```
```{r}
# remove "_new" term from canclation policy in order to reduce the combinations
pattern_removal <- function(x){
    return(str_remove(x, "_new"))
}
airbnb$Cancellation.Policy <- sapply(airbnb$Cancellation.Policy, pattern_removal)
```
Create new features¶
For each column we first find the all unique value exist, then we split the the specific column with a delimiter which is necessary (here ","). Next step we loop through the generated splited object and check each record for existing values. If the value exist we assign a 1 to that record with a new colun in airbnb dataframe

```{r}
pb <- txtProgressBar(min = 0, max = nrow(airbnb), style = 3)

temp = strsplit(as.character(airbnb$Host.Verifications),",")

for(j in 1:length(temp)){
    lst = temp[[j]]
    if('email' %in% lst){
        airbnb[j, 'email'] <- 1
    }
    if('reviews' %in% lst){
        airbnb[j, 'reviews'] <- 1
    }
    if('kba' %in% lst){
        airbnb[j, 'kba'] <- 1
    }
    if( 'facebook' %in% lst){
        airbnb[j, 'facebook'] <- 1
    }
    if( 'phone' %in% lst){
        airbnb[j, 'phone'] <- 1
    }
    if( 'government_id' %in% lst){
        airbnb[j, 'government_id'] <- 1
    }
    if( 'jumio' %in% lst){
        airbnb[j, 'jumio'] <- 1
    }
    if( 'work_email' %in% lst){
        airbnb[j, 'work_email'] <- 1
    }
    setTxtProgressBar(pb, j)
}

close(pb)
```

```{r}
pb <- txtProgressBar(min = 0, max = nrow(airbnb), style = 3)
temp = strsplit(as.character(airbnb$Features),",")
for(j in 1:length(temp)){
    lst = temp[[j]]
    if('Host Has Profile Pic' %in% lst){
        airbnb[j, 'Host Has Profile Pic'] <- 1
    }
    if('Host Identity Verified' %in% lst){
        airbnb[j, 'Host Identity Verified'] <- 1
    }
    if('Is Location Exact' %in% lst){
        airbnb[j, 'Is Location Exact'] <- 1
    }
    if( 'Instant Bookable' %in% lst){
        airbnb[j, 'Instant Bookable'] <- 1
    }
    if( 'Host Is Superhost' %in% lst){
        airbnb[j, 'Host Is Superhost'] <- 1
    }
    if( 'government_id' %in% lst){
        airbnb[j, 'government_id'] <- 1
    }
    if( 'Require Guest Phone Verification' %in% lst){
        airbnb[j, 'Require Guest Phone Verification'] <- 1
    }
    if( 'Require Guest Profile Picture' %in% lst){
        airbnb[j, 'Require Guest Profile Picture'] <- 1
    }
    if( 'Requires License' %in% lst){
        airbnb[j, 'Requires License'] <- 1
    }
    setTxtProgressBar(pb, j)
}

close(pb)
```

```{r}
#Replace missing/empty values on the generated columns to 0
airbnb[, 38:53][is.na(airbnb[, 38:53])] <- 0
write.csv(airbnb, 'airbnb_master_table.csv')
```
Working with missing values

```{r}
# Replace missing values with median of the column
for(i in 1:ncol(airbnb)){
    if(is.numeric(airbnb[,i])){
        airbnb[is.na(airbnb[,i]), i] <- median(airbnb[,i], na.rm = TRUE)
    }
}
nums <- unlist(lapply(airbnb, is.numeric))
cor(airbnb[, nums])
```
```{r}
corheat <- round(cor(airbnb[, nums]),2)
library(reshape2)
melted_corheat <- melt(corheat)
head(melted_corheat)
```
```{r}
library(ggplot2)
ggplot(data = melted_corheat, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile() + theme_classic() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + xlab('Numerical Variables') + ylab('Numerical Varibles')
```
```{r}
str(airbnb)
```
```{r}
by_country <- airbnb %>% group_by(Country) %>% summarise(mean_price = mean(Price, na.rm = TRUE)) 
by_country <- by_country[-c(1), ]
by_country <- by_country[order(by_country$mean_price),]
```
Price vs country
```{r}
library(ggplot2)
ggplot(data=by_country, aes(x=Country, y= mean_price)) + geom_bar(stat="identity", fill = 'black') + 
xlab('Country') + ylab('Mean Price') + theme_classic() + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
Amenities vs ratings
```{r}
library(magrittr)
#Selecting only the required columns
amenities_df <- airbnb %>%
  dplyr::select(ID,Amenities,`Review.Scores.Rating`,`Host Is Superhost`)
amenities_df <- na.omit(amenities_df)
```
```{r}
#changing the amenities column into different columns
library(data.table)
setDT(amenities_df)
amenities_df[, rn := seq_len(.N)]
long_amenities <- amenities_df[, strsplit(Amenities, ",", fixed = TRUE), by = rn]
```
```{r}
long_amenities[, V1 := stringr::str_trim(stringr::str_replace_all(V1, '["]', ""))]
amenities_sep <- dcast(long_amenities[!V1 %like% "^translation missing"], rn ~ V1, length, value.var = "rn", fill = 0)


amenities_df <- amenities_df[,c(1,3,4,5)]
final_amenities <- merge(amenities_sep, amenities_df, by = "rn")
```
```{r}
#Seperating highly rated listings
a_high <- final_amenities %>%
  filter(final_amenities$`Review.Scores.Rating` >=95) 
```


```{r}
a_high <- a_high[, !names(a_high) %in% c("rn", "Review.Scores.Rating","ID")] 
a_high_sum <- colSums(a_high)
a_high_df <-data.frame("Amenities"= colnames(a_high))
a_high_df$Sum <- a_high_sum
a_high_df <- a_high_df[order(a_high_df$Sum),]
a_high_df <- a_high_df %>%
  filter(a_high_df$Sum >= 2500)
```
```{r}
ggplot(a_high_df,aes(x=reorder(Amenities,Sum),y = Sum,fill = a))+geom_bar(stat = "identity", fill="steelblue")+coord_flip()+labs(x="Amenities",y="Number of listings",title="Highly rated listings")
```
```{r}
ggplot(airbnb, aes(x=Cleaning.Fee, y=Review.Scores.Cleanliness)) + geom_point() + labs(title = "Cleaning fee vs cleanliness score")
```
```{r}
plot <- summarise(group_by(airbnb,Cancellation.Policy,`Host Is Superhost`), count.plot = n())
plot$`Host Is Superhost`<-ifelse(plot$`Host Is Superhost`=="1",TRUE,FALSE)
plot <- plot[-1,]
ggplot(plot, aes(x=Cancellation.Policy,y = count.plot ,fill=`Host Is Superhost`)) + geom_bar(stat = "identity")
```
```{r}
plot2 <- summarise(group_by(airbnb,Country,Property.Type),count.plot2 = n())
plot2$Percentage <- (plot2$count.plot2 / sum(plot2$count.plot2))*100

ggplot(plot2,aes(x=Country,y=Percentage), fill= Property.Type)+geom_bar(stat="identity")
```
```{r}
airbnb$Superhost<- ifelse(airbnb$`Host Is Superhost`=="1",TRUE,FALSE)
ggplot(airbnb,aes(x= Host.Acceptance.Rate,y=Review.Scores.Rating))+geom_point(aes(colour=Superhost))+labs(title="What it takes to be a superhost")                                   
```
```{r}
library(alluvial)
filteredColumns <- airbnb %>% select(`Instant Bookable`,Cancellation.Policy, Room.Type,`Host Is Superhost`)
```